Planning with Hierarchical Temporal Memory for Deterministic Markov Decision Problem

### 2021-01-31 reports/v1_0/paper_eng/paper_eng.tex

In our work we present a simplified memory-based framework based on the ideas of HTM that can be used to model the environment dynamics. Great part of our contribution is that we also supply it with the planning algorithm that enables an agent to infer the optimal global policy. Some attempts to extend the HTM model have been made before \cite{Skrynnik2016,Daylidyonok2018,Nugamanov2020}, but the full implementation of the planning subsystem with consideration for reinforcement signal was not implemented in other works. We study applicability of the presented memory model and viability of the planning method in different maze environments.

@@

We introduced a novel memory-based method that combines learning the model of the environment with the planning algorithm to infer a local policy. We showed that under certain conditions (enough planning horizon) the resulting global policy is optimal or at least very close to it. We also found that in such cases, compared to DQN, our method learns with the comparable speed or slightly faster in terms of the number of episodes required. The planning algorithm with any non-zero planning horizon performs better than random strategy. However, for the small planning horizons and hard tasks results are sufficiently degraded and close to a pure random strategy.

@@

Our experiments showed us that the naive scaling of our method to environments with the larger state space is limited due to limitations for planning horizon increase. As for now we see two potential paths to remedy this problem. The first is to learn the model in a compact latent-space. The second is to make use of hierarchical approach by building a hierarchy of models and planners that operate in a different scales of time and space. Also the fact that the backing up strategy for a planner is purely random and not based on the agent's experience makes it a good target for further improvements too.

>>>>

In our work we present a simplified memory-based framework based on the ideas of HTM that can be used to model the environment dynamics. One of the main parts of our contribution is that we also supply it with the planning algorithm that enables an agent to infer the optimal global policy. Some attempts to extend the HTM model have been made before \cite{Skrynnik2016,Daylidyonok2018,Nugamanov2020}, but the full implementation of the planning subsystem with consideration for reinforcement signal was not implemented in other works. We study applicability of the presented memory model and viability of the planning method in different maze environments.

@@

We introduced a novel memory-based method that combines learning the model of the environment with the planning algorithm to infer a local policy. It is shown that under certain conditions (enough planning horizon) the resulting global policy is optimal or at least very close to it. We also found that in such cases, compared to DQN, our method learns with the comparable speed or slightly faster in terms of the number of episodes required. The planning algorithm with any non-zero planning horizon performs substantially better than random strategy. However, for the small planning horizons and hard tasks results are sufficiently degraded and close to a pure random strategy.

@@

Our experiments showed that the naive scaling of our method to environments with the larger state space is bounded due to limitations for planning horizon increase. As for now we see two potential paths to relieve the limits. One is to learn the model in a compact latent-space. Another is to make use of hierarchical approach by building a hierarchy of models and planners that operate in a different scales of time and space. Also the fact that the backing up strategy for a planner is purely random and not based on the agent's experience makes it a good target for further improvements too.
